# NGS filter files creation

Alrdy demultiplexed, no need fur ngs filters

## Primers

* 16s/C (412 pb):
    + PCR1-515F 5' CTTTCCCTACACGACGCTCTTCCGATCT GTGYCAGCMGCCGCGGT 3' (tag + primer) [tag is illumina adaptor?]
    + PCR1-909R 5' GGAGTTCAGACGTGTGCTCTTCCGATCT CCCCGYCAATTCMTTTRAGT 3' (tag + primer)
                                                
* 23s/A (410pb):
    + P23SrV_f1 5' CTTTCCCTACACGACGCTCTTCCGATCT GGACAGAAAGACCCTATGAA 3' (tag + primer)
    + P23SrV_r1 5' GGAGTTCAGACGTGTGCTCTTCCGATCT CAGCCTGTTATCCCTAGAG 3' (tag + primer)
    
* 18s/B (415pb / 200/-600) Tanabe et al. 2016 (stoeck et al. 2010):
    + TAReuk454FWD1: 5' CCAGCASCYGCGGTAATTCC 3' (primer)
    + TAReukREV3: 5' ACTTTCGTTCTTGATYRA 3' (primer)
    
degenerate nucleotides code:

--------------------
A |	Adenine |	A
C |	Cytosine |	C
G |	Guanine |	G
T |	Thymine (DNA) |	T
U |	Uracil (RNA) |	U
W |	Weak |	A/T
S |	Strong |	C/G
M |	Amino |	A/C
K |	Keto |	G/T
R |	Purine |	A/G
Y |	Pyrimidine |	C/T
B |	Not A |	C/G/T
D |	Not C |	A/G/T
H |	Not G |	A/C/T
V |	Not T |	A/C/G
N |	Any |	A/C/G/T

```{r}
nb_of_samples23s <- length(list.files("../data/sequencing/MAPP_23S/"))/2
nb_of_samples18s <- length(list.files("../data/sequencing/MAPP_18S/"))/2
nb_of_samples16s <- length(list.files("../data/sequencing/MAPP_16S/"))/2
```


```{r}
ngs_16s <- data.frame(sample_id=unique(gsub("VJ-","",(gsub("_.*gz","",list.files("../data/sequencing/MAPP_16S/"))))),
                      tags=unique(stringr::str_extract(list.files("../data/sequencing/MAPP_16S/"),"(?<=_)[A-Z]+(?=-)")),
                      seq_prime_f="GTGYCAGCMGCCGCGGT",
                      seq_prime_r="CCCCGYCAATTCMTTTRAGT")


unique(gsub("VJ-","",(gsub("_.*gz","",list.files("../data/sequencing/MAPP_16S/")))))
unique(stringr::str_extract(list.files("../data/sequencing/MAPP_16S/"),"(?<=_)[A-Z]+(?=-)"))

duplicated((stringr::str_extract(list.files("../data/sequencing/MAPP_16S/"),"(?<=_)[A-Z]+(?=-)")))
stringr::str_extract(list.files("../data/sequencing/MAPP_16S/"),"(?<=_)[A-Z]+(?=-)")
```


## Frogs news

https://frogs.toulouse.inra.fr/news/FROGSNews-June2023.html#FROGS-v410-is-available
